---
title: "Red Deer sociality and fitness"
author: "Sebastian Sosa"
format: html
editor: visual
---

# Dependencies

```{r}
#| message: false
#| warning: false
library(ANTs)
library(lme4)
library(lmerTest)
library(sjPlot)
library(ggplot2)
library(MASS)
library(asnipe)
library(survival)
library(survminer)
library(rptR)
library(PerformanceAnalytics)
library(gtsummary)
library(coxme)
theme_set(
  theme_minimal(base_size = 14) +  # base_size sets most text sizes
    theme(
      plot.title = element_text(size = 18, face = "bold"),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 12),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 12)
    )
)
update_geom_defaults("point", list(size = 16))
```


# Analysis
## MRQAP
```{r}
load("MRQAP_data.RData")
mrqap.mum = sample = NULL

# for each networks
for (a in 1:length(original.network)) {
  # Get only individual present in all data sets
  ids = Reduce(intersect, list(colnames(original.network[[a]]),
                               colnames(ped.mum[[a]]),
                               colnames(mat.HRO[[a]]),
                               colnames(mat.spde[[a]]),
                               colnames(mat.spdeX[[a]]),
                               colnames(mat.spdeY[[a]])))
  
  if(length(ids) <2){
    print(c("to few individuals in ", names(original.network)[a]))
    mrqap.mum[[a]] = NULL
    next
  }
  
  
  net = original.network[[a]][colnames(original.network [[a]]) %in% ids, colnames(original.network [[a]]) %in% ids]
  
  net = net[order(colnames(net)),order(colnames(net))]
  
  ped = ped.mum[[a]][colnames(ped.mum[[a]]) %in% ids, colnames(ped.mum[[a]]) %in% ids]
  ped = ped[order(colnames(ped)),order(colnames(ped))]
  
  hro = mat.HRO[[a]][colnames(mat.HRO[[a]]) %in% ids, colnames(mat.HRO[[a]]) %in% ids]
  hro = hro[order(colnames(hro)),order(colnames(hro))]
  
  spde = mat.spde[[a]][colnames(mat.spde[[a]]) %in% ids, colnames(mat.spde[[a]]) %in% ids]
  
  spdeX = mat.spdeX[[a]][colnames(mat.spdeX[[a]]) %in% ids, colnames(mat.spdeX[[a]]) %in% ids]
  
  spdeY = mat.spdeY[[a]][colnames(mat.spdeY[[a]]) %in% ids, colnames(mat.spdeY[[a]]) %in% ids]
  
  ped = ped[order(colnames(ped)),order(colnames(ped))]
  
  sample = c(sample,ncol(net))
  if(any(c(is.null(ncol(net)),
           is.null(ncol(ped)),
           is.null(ncol(hro )),
           is.null(ncol(spde))))){next()}
  
  if(any(c(all(net == 0),
           all(ped == 0),
           all(hro == 0),
           all(spde == 0)))){next()}
  # MRQAP
  mrqap.mum[[a]] = asnipe::mrqap.dsp(net ~ ped + hro + spdeX + spdeY)
  names(mrqap.mum)[a] = names(original.network)[a]
}


names(mrqap.mum)= names(original.network)
mrqap.mum = mrqap.mum[-1]
mrqap.p = lapply(mrqap.mum, function(x){
  c = x$P.values[2]
})
mrqap.p = unlist(mrqap.p)
names(mrqap.p) = names(mrqap.mum)

mrqap.coef = lapply(mrqap.mum, function(x){
  x$coefficients[2]
})
mrqap.coef= unlist(mrqap.coef)
names(mrqap.coef) = names(mrqap.mum)

t = ifelse(mrqap.p < 0.05, "singificant", "non singificant")
```
### Figure 2
```{r}
ggplot(data = NULL, aes(y = mrqap.coef, x= names(mrqap.coef)))+geom_point(size = 8)+geom_line()+theme(legend.position = "none")+ylab("Regression coefficients (red =  significant value)")+xlab("Year")
```

## Repetability
```{r}
#| message: false
#| warning: false
annual=read.csv('annual_data.csv')
rep = rpt(formula = strength ~ (1|id), grname = "id", data = annual, datatype = "Gaussian", npermut = 10000)
rep
```

## LRS
```{r}
repro=read.csv('scaled_life_time_data.csv') # scaled data for analysis
repro2=read.csv('life_time_data.csv') # non scale data for metrics descriptions
```

### Measures summary
```{r}
tbl_summary(
  repro2[,colnames(repro2) %in% c('sum_degree', 'sum_strength',  'sum_eigen')],
  missing = "no" # Don't show missing values
) %>%
bold_labels() 

```

### Network metrics correlations (Appendix 1)
```{r}
chart.Correlation(repro[,colnames(repro) %in% c("sum_eigen", "sum_strength", "sum_degree")], histogram=TRUE, pch=19)
```
### LBS poisson model
```{r}
m1=glm(LBS~sum_strength+lifespan+mean_matSize+mean_E+mean_N+mean_HRO, family = poisson(), data = repro)
print(car::vif(m1)) # VIF
```

### Table 2
```{r}
tab_model(m1, transform = NULL, show.se = TRUE, digits = 3, pred.labels = c('Intercept', 'Life Time Strength','Lifespan', 'Matriline Size', 'Annual mean East', 'Annual mean North', 'Annual mean HRO'))
```

### Figure 3
```{r}
p1 = plot_model(m1, type = 'pred', terms ='lifespan', show.data = T)+xlab('Lifespan')
p2 = plot_model(m1, type = 'pred', terms ='sum_strength', show.data = T)+xlab('Life Time Strength')
ggarrange(p1,p2, ncol = 2, nrow = 1)
```


### LRS poisson model
```{r}
m2=glm(LRS~sum_strength+lifespan+mean_matSize+mean_E+mean_N+mean_HRO, family = poisson(), data = repro)
print(car::vif(m2)) # VIF
```

### Table 3
```{r}
tab_model(m2, transform = NULL, show.se = TRUE, digits = 3, pred.labels = c('Intercept', 'Life Time Strength','Lifespan', 'Matriline Size', 'Annual mean East', 'Annual mean North', 'Annual mean HRO'))
```

### Figure 4
```{r}
p3 = plot_model(m2, type = 'pred', terms ='lifespan', show.data = T)+xlab('Lifespan')
p4 = plot_model(m2, type = 'pred', terms ='sum_strength', show.data = T)+xlab('Life Time Strength')
ggarrange(p3,p4, ncol = 2, nrow = 1)
```

## Survival analysis
### Data
```{r}
surv = read.csv(file='scale_annual.csv')
```

### Proportional Hazards Regression Model 
```{r}
surv_obj <- Surv(surv$start, surv$end, surv$death)
```


```{r}
model_coxme <- coxme(surv_obj ~ strength + matSize + HRO2 + LifetimeE + LifetimeN + (1 | mat/id),data = surv,  na.action = na.omit)
summary(model_coxme)
```



### Model diagnostic
#### Test the Proportional Hazards Assumption of a Cox Regression
```{r}
ph_test <- cox.zph(model_coxme)
print(ph_test)
```

#### residual heterodasiticity
```{r}
rows_omitted <- model_coxme$na.action
if (!is.null(rows_omitted)) {
  model_data <- surv[-rows_omitted, ]
} else {
  model_data <- surv
}
mresid <- residuals(model_coxme, type = "martingale")


# For strength
plot(model_data$strength, mresid,
     xlab = "Strength", ylab = "Martingale Residuals")
lines(lowess(model_data$strength, mresid), col = "red")
abline(h = 0, lty = 2)

# For LifetimeE
plot(model_data$LifetimeE, mresid,
     xlab = "LifetimeE", ylab = "Martingale Residuals",
     main = "Martingale Residuals vs. LifetimeE")
lines(lowess(model_data$LifetimeE, mresid), col = "red")
abline(h = 0, lty = 2)

# For LifetimeN
plot(model_data$LifetimeN, mresid,
     xlab = "LifetimeN", ylab = "Martingale Residuals",
     main = "Martingale Residuals vs. LifetimeN")
lines(lowess(model_data$LifetimeN, mresid), col = "red")
abline(h = 0, lty = 2)

```

### Figure 5 
```{r}
summa = summary(surv$strength)
surv$strength2 = NULL
surv$strength2 = ifelse(surv$strength >= summa[[5]], "3rd Qu.", surv$strength)
surv$strength2 = ifelse(surv$strength >= summa[[2]] & surv$strength < summa[[5]], "mean", surv$strength2)
surv$strength2 = ifelse(surv$strength < summa[[2]], "1st Qu.", surv$strength2)
survfit(Surv(surv$start, surv$end, surv$death)~strength2 , data=surv)
ggsurvplot(
  survfit(
    Surv(surv$start, surv$end, surv$death)~strength2, data=surv),
  data = surv, 
  fun = "event", 
  conf.int = TRUE, 
  risk.table = TRUE, 
  risk.table.col="strata", 
  ggtheme =  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 18),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 18)
  ))+
  xlab("Age")   

```

## Calves' survival during the first year
### Data
```{r}
calves = read.csv(file='scaled_calves_survival.csv')
```
### Binomial model
```{r}
m3 = glmer(ego.survived.frist.year ~  mom.strength +  matSize + HRO2 + LifetimeE + LifetimeN +(1|ego.birth)  + (1|ego.mom) + (1|matriline), family = binomial(), data = calves)
tab_model(m3, transform = NULL, show.se = TRUE, digits = 3, pred.labels = c('Intercept', 'Mother strength','Matriline size', 'HRO', 'East', 'North'))
```
### Figure
```{r}
t = ggeffects::ggpredict(m3, terms = "LifetimeE [all]")

p5 = ggplot(data = calves, aes(x = LifetimeE, y = ego.survived.frist.year)) +
  geom_histogram(data = calves[calves$ego.survived.frist.year ==0, ], aes(x = LifetimeE, y = stat(count)/1000, fill=factor(ego.survived.frist.year)), bins = 15, na.rm = TRUE) +
  geom_histogram(data = calves[calves$ego.survived.frist.year == 1, ], aes(x = LifetimeE, y = -1*stat(count/1000), fill=factor(ego.survived.frist.year)), bins = 15, na.rm = TRUE, position = position_nudge(y = 1))+
  theme(legend.position = "none")+
  geom_ribbon(data= t, aes(x = x,ymin= conf.low, ymax= conf.high, y= NULL), fill= 'grey80')+
  geom_line(data = t, aes(x = x , y = predicted ))+
  ylab("First year survival probability")+
  xlab("East")
p5

```
